### 2015-06-06 chapters/pitfalls.tex
correction some latex errors

Unicode is a character encoding standard, but these characters of course
actually appear inside some kind of file. The most basic Unicode-based file
format is pure line-based text, i.e.~strings of Unicode-encoded characters
separated by line breaks (note that these line breaks are what for most people
intuitively corresponds to paragraph breaks). Unfortunately, even within this
apparently basic setting there exist a multitude of variants. In general, these
different possibilities are well-understood in the software industry, and
nowadays they normally do not lead to any problems for the end user. However,
there are some situations in which a user is suddenly confronted with cryptic
questions in the user interface involving abbreviations like LF, CR, BE, LE or
BOM. Most prominently this occurs with exporting or importing data in several
software applications from Microsoft. Basically, there are two different issues
involved. First, the encoding of line breaks and, second, the encoding of the
Unicode characters into \textsc{code units} and the related issue of
\textsc{endianness}.

@@

The issue with \textsc{line breaks} originated with the instructions necessary
to direct a printing head to a new line. This involves two movements, known as
\textsc{carriage return} (CR, returning the printing head to the start of the
line on the page) and \textsc{line feed} (LF, moving the printing head to the
next line on the page). Physically, these are two different events, but
conceptually together they form one action. In the history of computing, various
encodings of line breaks have been used (e.g.~CR+LF, LF+CR, only LF, or only CR.
Currently, all Unix and Unix-derived systems use only LF as code for a
line break, while software from Microsoft still uses a combination of
CR and LF. Today, most software applications recognize both
options, and are able to deal with either encoding of line breaks (until rather
recently this was not the case, and using the wrong line breaks could lead to
unexpected errors). Our impression is that there is a strong tendency in
software development to standardize on the simpler \textsc{only LF}
encoding for line breaks, and we suggest that everybody use this encoding
whenever possible.

@@

The issue with \textsc{code units} stems from the question how to separate a
stream of binary ones and zero, i.e.~bits, into chunks representing Unicode
characters. A code unit is the sequence of bits used to encode a single
character in an encoding. Depending on different use cases, the Unicode Standard
offers three different approaches, called UTF-32, UTF-16 and UTF-8.\footnote{The
letters UTF stand for either \textsc{Unicode Transformation Format} or
\textsc{Universal Coded Character Set Transformation Format}, but the notion of
\textit{transformation} is a legacy notion that does not have meaning anymore.
Nevertheless, the designation \textsc{UTF} (in capitals) has become an official
standard designation, but should probably best be read as simply Unicode
Format.} The details of this issue is extensively explained in section 2.5 of
the Unicode Core Specification \cite{Unicode2014}.

>>>>

Unicode is a character encoding standard, but these characters of course
actually appear inside some kind of file. The most basic Unicode-based file
format is pure line-based text, i.e.~strings of Unicode-encoded characters
separated by line breaks (note that these line breaks are what for most people
intuitively corresponds to paragraph breaks). Unfortunately, even within this
apparently basic setting there exist a multitude of variants. In general, these
different possibilities are well-understood in the software industry, and
nowadays they normally do not lead to any problems for the end user. However,
there are some situations in which a user is suddenly confronted with cryptic
questions in the user interface involving abbreviations like LF, CR, BE, LE or
BOM.\@ Most prominently this occurs with exporting or importing data in several
software applications from Microsoft. Basically, there are two different issues
involved. First, the encoding of line breaks and, second, the encoding of the
Unicode characters into \textsc{code units} and the related issue of
\textsc{endianness}.

@@

The issue with \textsc{line breaks} originated with the instructions necessary
to direct a printing head to a new line. This involves two movements, known as
\textsc{carriage return} (CR, returning the printing head to the start of the
line on the page) and \textsc{line feed} (LF, moving the printing head to the
next line on the page). Physically, these are two different events, but
conceptually together they form one action. In the history of computing, various
encodings of line breaks have been used (e.g.~CR+LF, LF+CR, only LF, or only
CR). Currently, all Unix and Unix-derived systems use only LF as code for a
line break, while software from Microsoft still uses a combination of CR and
LF.\@ Today, most software applications recognize both options, and are able to
deal with either encoding of line breaks (until rather recently this was not the
case, and using the wrong line breaks could lead to unexpected errors). Our
impression is that there is a strong tendency in software development to
standardize on the simpler \textsc{only LF} encoding for line breaks, and we
suggest that everybody use this encoding whenever possible.

@@

The issue with \textsc{code units} stems from the question how to separate a
stream of binary ones and zero, i.e.~bits, into chunks representing Unicode
characters. A code unit is the sequence of bits used to encode a single
character in an encoding. Depending on different use cases, the Unicode Standard
offers three different approaches, called UTF-32, UTF-16 and UTF-8.\footnote{The
letters UTF stand for either \textsc{Unicode Transformation Format} or
\textsc{Universal Coded Character Set Transformation Format}, but the notion of
\textit{transformation} is a legacy notion that does not have meaning anymore.
Nevertheless, the designation \textsc{UTF} (in capitals) has become an official
standard designation, but should probably best be read as simply Unicode
Format.} The details of this issue is extensively explained in section 2.5 of
the Unicode Core Specification~\cite{Unicode2014}.
